Malacology: A Programmable Storage System

### 2016-10-03 paper/paper.tex
paper: jeff's section 4 comments; start decoupling Ceph
Signed-off-by: Michael Sevilla <michael.sevilla@hpe.com>

\usepackage{graphicx}
\usepackage[font=footnotesize,labelfont=bf]{caption}
\usepackage{pstricks}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Malacology: A Programmable Storage System Built on Ceph},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
%\setlength{\parindent}{0pt}
%\setlength{\parskip}{6pt plus 2pt minus 1pt}
%\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

\title{Malacology: A Programmable Storage System Built on Ceph}
\author{Paper \# 123}
\date{}
\begin{document}

@@

\begin{abstract}
Storage systems are caught between rapidly changing data processing systems and
the increasing speed of storage devices. This puts tremendous pressure on
storage systems to adapt both in terms of their interfaces and their
performance. But adapting storage systems can be difficult because unprincipled
changes might jeopardize years of code-hardening and performance optimization
efforts that were necessary for users to entrust their data to the storage
system.  We introduce  Malacology, a prototype programmable storage system to
explore how existing abstractions of common services found in storage systems
can be leveraged to address new data processing systems and the increasing
speed of storage devices. This approach allows unprecedented flexibility for
storage systems to evolve without sacrificing the robustness of its
code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by constructing two services out of existing abstractions: a
file system metadata load balancer and a high-performance distributed shared-log
that leverages flash devices. The evaluation demonstrates that our services 
inherit desirable qualities of Ceph, including the ability to make trade-offs 
for lease capabilities, balance load (2\(\times\) improvement), propagate interfaces 
(average of 222ms) and recover (15 second down time).  
\end{abstract}

@@

\begin{figure}[tb]
\centering
\includegraphics{figures/overview.png}
\caption{A Ceph installation consists of a cluster running mostly
storage daemons (OSDs) to ensure data durability, a few monitor daemons
(MONs) to reach consensus of the cluster state and to ensure its
consistent versioning, and a few metadata service daemons (MDSs) for
scalable metadata service. Ceph's APIs include object, block, file
abstractions. Malacology enables the programmability of these
abstractions and services (indicated by bold arrows). To illustrate Malacology we implement two
services on Ceph, ZLog and Mantle, by programming Ceph's durability,
consistent versioning, consensus, and metadata subsystems.
\label{fig:overview}}
\end{figure}

@@

We will now discuss the subsystems Ceph uses to provide a general-purpose
storage system and how Malacology makes them programmable.

@@

\subsection{Mantle: A Programmable Metadata Load Balancer}
\label{sec:mantle}

Mantle is a programmable metadata
balancer for Ceph that separates the metadata balancing policies from their
mechanisms. Administrators inject code to change how the metadata cluster
distributes metadata. In ~\cite{sevilla:sc15-mantle} the authors showed how 
using Mantle, one can implement a single node metadata service, a distributed 
metadata services with hashing, and a distributed metadata service with dynamic 
subtree partitioning.

\subsection{Re-Implementation}

The Ceph team wants to merge Mantle because the scriptability is useful for
debugging, controlling the metadata balancer, and examining trade-offs for
different balancers. Unfortunately, this research quality system is not as
robust as Ceph and the Ceph team wants more safety, durability, and consistency
for the new services.  The API for Malacology is the same as the one exposed
in~\cite{sevilla:sc15-mantle} but the hooks are put in different places.

\subsection{Policy Versioning}

For Mantle, we added new monitor command to change the metadata load balancers:

\noindent \texttt{ mds\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}
%\noindent \texttt{ceph\ osd\ pool\ set-class\ \textless{}pool\textgreater{}\ \textless{}class\textgreater{}\ \textless{}script\textgreater{}}

This command assigns the current version of the metadata load balancer. In our case, it is simply the name of the balancer. The MONs take the version and store it in the MDS map. Similar to the PG map or MON maps, the MDS map has all the information about the topology and state of the MDS cluster. Leveraging the MONs helps ease the burden on the MDS to propagate the policy and ensure that everyone agrees.

There are Ceph components that are obstacles to this approach, namely the size of the  MDS map, API changes, and security. The Ceph engineers want to avoid using the maps to exchange information; the maps should be lightweight and fast because the MONs need to be able to version MDS cluster state quickly. API changes can also be a burden, especially for the MDS map, since all clients need to have the same version of the CephFS library in order to properly interact with CephFS. Finally, Ceph tries to avoid injecting strings into the cluster because they are insecure and buggy.

\subsection{Making Policies Durable}

While storing the code snippets that define the policies in the MDS map is 
possible, it is not recommended as it is an expensive operation (a PAXOS voting 
round). Thus, Malacology only stores policies IDs which correspond to an object 
in a system-maintained pool. When MDSs read a policy, they first check latest 
version from the MDS map, and then dereference the pointer by reading the 
corresponding object in RADOS.

>>>>

\usepackage{graphicx}
\usepackage[font=footnotesize,labelfont=bf]{caption}
\usepackage{pstricks}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Malacology: A Programmable Storage System},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
%\setlength{\parindent}{0pt}
%\setlength{\parskip}{6pt plus 2pt minus 1pt}
%\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

\title{Malacology: A Programmable Storage System}
\author{Paper \# 123}
\date{}
\begin{document}

@@

\begin{abstract}
Storage systems are caught between rapidly changing data processing systems and
the increasing speed of storage devices. This puts tremendous pressure on
storage systems to adapt both in terms of their interfaces and their
performance. But adapting storage systems can be difficult because unprincipled
changes might jeopardize years of code-hardening and performance optimization
efforts that were necessary for users to entrust their data to the storage
system.  We introduce  Malacology, a prototype programmable storage system to
explore how existing abstractions of common services found in storage systems
can be leveraged to address new data processing systems and the increasing
speed of storage devices. This approach allows unprecedented flexibility for
storage systems to evolve without sacrificing the robustness of its
code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by constructing two services out of existing abstractions: a
file system metadata load balancer and a high-performance distributed shared-log
that leverages flash devices. The evaluation demonstrates that our services 
inherit desirable qualities of the back-end storage system, including the ability to make trade-offs 
for lease capabilities, balance load (2\(\times\) improvement), propagate interfaces 
(average of 222ms) and recover (15 second down time).  
\end{abstract}

@@

\begin{figure}[tb]
\centering
\includegraphics{figures/overview.png}

\caption{Scalable storage systems consist of object storage daemons (OSD) which
store data fragments, monitor daemons (MON) that maintain cluster state, and
service-specific daemons such as file system metadata servers (MDS). Malacology
enables the programmability of internal abstractions and services (indicated by
bold arrows) in order to leverage the back-end system's durability, consistent
versioning, consensus, and metadata subsystems.  With Malacology, we add two
new services, ZLog and Mantle, that sit alongside the traditional user-facing
APIs (file, block, and object) in the storage stack.  \label{fig:overview}}
\end{figure}

@@

We will now discuss the common subsystems used to manage
storage system and how Malacology makes them programmable.

@@

\subsection{Mantle: Programmable Load Balancer}
\label{sec:mantle}

Mantle is a programmable load balancer that separates the metadata balancing
policies from their mechanisms. Administrators inject code to change how the
metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
authors showed how using Mantle, one can implement a single node metadata
service, a distributed metadata services with hashing, and a distributed
metadata service with dynamic subtree partitioning.

The original implementation was ``hard-coded" into Ceph and lacked robustness
(lacking things such as versioning, durability, or reliable policy
distribution).  Re-implemented using Malacology, Mantle now enjoys (1) the
versioning of MONs and (2) the durability and distribution provided by RADOS.
The version of the load balancer, which is stored in the MDS map, corresponds
to an object name in RADOS that houses the load balancer code.

\subsubsection{Versioning Balancer Policies}

Ensuring that the version of the current load balancer is consistent across the
MDS cluster was ignored in the original implementation. The user had to set the
version on each individual MDS and it was trivial to make the version
inconsistent across MDS nodes.

With Malacology, Mantle stores the version of the current load balancer in the
MDS map.  Similar to the PG map or MON maps, the MDS map has all the
information about the topology and state of the MDS cluster.  Recall that the
MONs propogate these maps across the cluster and ensure that the map versions
are consistent with PAXOS.  While storing the code snippets that define the
policies in the MDS map is possible, it is not recommended as it is an
expensive operation (a PAXOS voting round).

The user changes the version of the load balancer using a new CLI command:

\noindent \texttt{ fs\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}

%There are Ceph components that are obstacles to this approach, namely the size
%of the  MDS map, API changes, and security. The Ceph engineers want to avoid
%using the maps to exchange information; the maps should be lightweight and fast
%because the MONs need to be able to version MDS cluster state quickly. API
%changes can also be a burden, especially for the MDS map, since all clients
%need to have the same version of the CephFS library in order to properly
%interact with CephFS. Finally, Ceph tries to avoid injecting strings into the
%cluster because they are insecure and buggy.

\subsubsection{Making Balancer Policies Durable}

The load balancer version described above corresponds to the name of an object
in RADOS that holds the actual Lua balancing code. The objects containing load
balancer code are stored in a system-maintained group of objects (i.e. a pool).
When MDS nodes start balancing load, they first check the latest version from
the MDS map and compare it to the balancer they have loaded. If the version has
changed, they dereference the pointer to the balancer version by reading the
corresponding object in RADOS. This is in contrast to the original balancer
which stored load balancer code on the local file system -- a technique which
is unreliable and results in inconsistencies if any MDS accidently modifies the
code.

The balancer pulls the Lua code from RADOS synchronously; asynchronous reads
are not possible because of the architecture of the MDS. The synchronous
behavior is not the default behavior for RADOS operations, so we achieve this
with a timeout: if the asynchronous read does not come back within half the
balancing tick interval the operation is cancelled and a Connection Timeout
error is returned. By default, the balancing tick interval is 10 seconds, so
Mantle will use a 5 second second timeout.

This design allows Mantle to immediately return an error if anything
RADOS-related goes wrong.  We use this implementation because we do not want to
do a blocking OSD read from inside the global MDS lock. Doing so would bring
down the MDS cluster if any of the OSDs are not responsive.

Storing the balancers in RADOS is only possible because we use Lua as the
language for writing balancer code. If we used a language that needs to be
compiled, like the C++ object classes in the OSD, we would need to ensure
binary compatibility, which is complicated by different operating systems,
distributions, and compilers.

\subsubsection{Logging, Debugging, and Warnings}

In the original implementation, Mantle would log all errors, warnings, and
debug messages to the local log on each MDS. To get the simplest status
messages or to debug problems, the user would have to log into each MDS
individually, look at the logs, and reason about their causality.

With Malacology, Mantle re-usues the centralized logging features of the MON
subsystem. Important errors, warnings, and info messages are sent with the MON
beacon subsystem and appear in the monitor cluster log so instead of users
going to each node, they can watch messages appear at the monitor. Messages are
used discretly, so as not to spam the monitor with frivolous debugging but the
large events, like balancer version changes or failed subsystems show up in the
centralized log.
